---
title: "Wine Quality ML Regression Models"
author: "Josh Bicer"
date: "2023-06-01"
output:
  pdf_document: default
  html_document: default
---

```{r include=FALSE}
library(knitr)
library(triangle)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(scipen = 6)
```
### Part 1: Install libraries and Read Data
```{r}

# Install required packages and libraries
library(ggplot2) 
library(MASS) 
library(pscl) 
library(dplyr) 
library(readr)
library(corrplot)
library(zoo)
library(Amelia)
library(psych)
library(car)
library(glm2)
library(leaps)

# Read in data from CSV file
wine = read.csv("wine_training.csv")
wine_test = read.csv("wine_testing.csv")
```
### Part 2: Data Exploration
This section explores the wine quality dataset to gain a better understanding of the data used for the model. This includes histogram and boxplot visualizations for the target variable and all applicable input variables. Correlation plots and summaries of the dataset are included below.
```{r}
# Review data frame summary
summary(wine)
# Review structure of the data frame
str(wine)

# Histogram of target variable for Star Rating
# Rating is on a scale between 0 - 10
ggplot(data=wine, aes(TARGET)) + 
  geom_histogram(binwidth =1, 
                 col="BLUE", 
                 aes(fill=..count..))+
  scale_fill_gradient("Count", low = "blue", high = "red") +
  ggtitle("Histogram of Wine Quality Rating") +
  xlab("Rating") +
  ylab("Frequency")

zero_count = length(wine$TARGET[wine$TARGET == 0])
print(paste('The number of zeros in Target Variable is: ', zero_count))

# Examine correlation among variables
wine_clean = na.omit(wine)
cor(wine_clean[sapply(wine_clean, is.numeric)])
# Plot correlation
corrplot(cor(wine_clean), method = "square")

### Explore histograms and boxplots of independent variables
# Fixed Acidity and Volatile Acidity
par(mfrow=c(2,2))
hist(wine$FixedAcidity, col = "royalblue", xlab = "FixedAcidity", main = "Histogram of FixedAcidity")
hist(wine$VolatileAcidity, col = "red", xlab = "VolatileAcidity", main = "Histogram of VolatileAcidity")
boxplot(wine$FixedAcidity, col = "skyblue", main = "Boxplot of FixedAcidity")
boxplot(wine$VolatileAcidity, col = "orangered", main = "Boxplot of VolatileAcidity")
par(mfrow=c(1,1))

# Citric Acid and Residual Sugar
par(mfrow=c(2,2))
hist(wine$CitricAcid, col = "royalblue", xlab = "CitricAcid", main = "Histogram of CitricAcid")
hist(wine$ResidualSugar, col = "red", xlab = "ResidualSugar", main = "Histogram of ResidualSugar")
boxplot(wine$CitricAcid, col = "skyblue", main = "Boxplot of CitricAcid")
boxplot(wine$ResidualSugar, col = "orangered", main = "Boxplot of ResidualSugar")
par(mfrow=c(1,1))

# Chlorides and Free Sulfur Dioxide
par(mfrow=c(2,2))
hist(wine$Chlorides, col = "royalblue", xlab = "Chlorides", main = "Histogram of Chlorides")
hist(wine$FreeSulfurDioxide, col = "red", xlab = "FreeSulfurDioxide", main = "Histogram of FreeSulfurDioxide")
boxplot(wine$Chlorides, col = "skyblue", main = "Boxplot of Chlorides")
boxplot(wine$FreeSulfurDioxide, col = "orangered", main = "Boxplot of FreeSulfurDioxide")
par(mfrow=c(1,1))

# Total Sulfur Dioxide and Density
par(mfrow=c(2,2))
hist(wine$TotalSulfurDioxide, col = "royalblue", xlab = "TotalSulfurDioxide", main = "Histogram of TotalSulfurDioxide")
hist(wine$Density, col = "red", xlab = "Density", main = "Histogram of Density")
boxplot(wine$TotalSulfurDioxide, col = "skyblue", main = "Boxplot of TotalSulfurDioxide")
boxplot(wine$Density, col = "orangered", main = "Boxplot of Density")
par(mfrow=c(1,1))

# pH Level and Alocohol Content
par(mfrow=c(2,2))
hist(wine$pH, col = "royalblue", xlab = "pH", main = "Histogram of pH")
hist(wine$Alcohol, col = "red", xlab = "Alcohol", main = "Histogram of Alcohol")
boxplot(wine$pH, col = "skyblue", main = "Boxplot of pH")
boxplot(wine$Alcohol, col = "orangered", main = "Boxplot of Alcohol")
par(mfrow=c(1,1))

# Label appeal and STARS
par(mfrow=c(2,2))
hist(wine$LabelAppeal, col = "royalblue", xlab = "LabelAppeal", main = "Histogram of LabelAppeal")
hist(wine$STARS, col = "red", xlab = "STARS", main = "Histogram of STARS")
boxplot(wine$LabelAppeal, col = "skyblue", main = "Boxplot of LabelAppeal")
boxplot(wine$STARS, col = "orangered", main = "Boxplot of STARS")
par(mfrow=c(1,1))
```
### Part 3: Data Preparation
This section is used to clean the data to prepare for use in the machine learning models. All data cleaning steps are performed on both the training and testing datasets for consistency. Any missing values in the dataset are replaced with the median value and any significant outliers are replaced with the 1% or 99% value. Additionally, all original data is preserved and any imputed values in the model use the _IMP label to differentiate. 
```{r}
### Training Data Set
# Create imputed variables for each column
wine$FixedAcidity_IMP = wine$FixedAcidity
wine$VolatileAcidity_IMP = wine$VolatileAcidity
wine$CitricAcid_IMP = wine$CitricAcid
wine$ResidualSugar_IMP = wine$ResidualSugar
wine$Chlorides_IMP = wine$Chlorides
wine$FreeSulfurDioxide_IMP = wine$FreeSulfurDioxide
wine$TotalSulfurDioxide_IMP = wine$TotalSulfurDioxide
wine$Density_IMP = wine$Density
wine$pH_IMP = wine$pH
wine$Sulphates_IMP = wine$Sulphates
wine$Alcohol_IMP = wine$Alcohol
wine$LabelAppeal_IMP = wine$LabelAppeal
wine$AcidIndex_IMP = wine$AcidIndex
wine$STARS_IMP = wine$STARS

### Replace missing NA values with median and outliers with 1st or 99th percentile
### Create FLAG variables where replaced.
#ResidualSugar
wine$ResidualSugar_FLAG = as.factor(ifelse(wine$ResidualSugar_IMP < quantile(wine$ResidualSugar_IMP, c(.01), na.rm=TRUE)
                                           | wine$ResidualSugar_IMP > quantile(wine$ResidualSugar_IMP,c(.99), na.rm=TRUE)
                                           | is.na(wine$ResidualSugar_IMP),1,0))
wine$ResidualSugar_IMP[is.na(wine$ResidualSugar_IMP)] = median(wine$ResidualSugar_IMP, na.rm=TRUE)
wine$ResidualSugar_IMP = as.numeric(ifelse(wine$ResidualSugar_IMP < quantile(wine$ResidualSugar_IMP, c(.01)),
                                           quantile(wine$ResidualSugar_IMP,c(.01)), wine$ResidualSugar_IMP))
wine$ResidualSugar_IMP = as.numeric(ifelse(wine$ResidualSugar_IMP > quantile(wine$ResidualSugar_IMP, c(.99)),
                                           quantile(wine$ResidualSugar_IMP,c(.99)), wine$ResidualSugar_IMP))

#Chlorides
wine$Chlorides_FLAG = as.factor(ifelse(is.na(wine$Chlorides),1,0))
wine$Chlorides_IMP[is.na(wine$Chlorides_IMP)] = median(wine$Chlorides, na.rm = TRUE)

#FreeSulfurDioxide
wine$FreeSulfurDioxide_FLAG = as.factor(ifelse(wine$FreeSulfurDioxide_IMP < quantile(wine$FreeSulfurDioxide_IMP, c(.01), na.rm=TRUE)
                                           | wine$FreeSulfurDioxide_IMP > quantile(wine$FreeSulfurDioxide_IMP,c(.99), na.rm=TRUE)
                                           | is.na(wine$FreeSulfurDioxide_IMP),1,0))
wine$FreeSulfurDioxide_IMP[is.na(wine$FreeSulfurDioxide_IMP)] = median(wine$FreeSulfurDioxide_IMP, na.rm=TRUE)
wine$FreeSulfurDioxide_IMP = as.numeric(ifelse(wine$FreeSulfurDioxide_IMP < quantile(wine$FreeSulfurDioxide_IMP, c(.01)),
                                           quantile(wine$FreeSulfurDioxide_IMP,c(.01)), wine$FreeSulfurDioxide_IMP))
wine$FreeSulfurDioxide_IMP = as.numeric(ifelse(wine$FreeSulfurDioxide_IMP > quantile(wine$FreeSulfurDioxide_IMP, c(.99)),
                                           quantile(wine$FreeSulfurDioxide_IMP,c(.99)), wine$FreeSulfurDioxide_IMP))

#TotalSulfurDioxide
wine$TotalSulfurDioxide_FLAG = as.factor(ifelse(wine$TotalSulfurDioxide_IMP < quantile(wine$TotalSulfurDioxide_IMP, c(.01), na.rm=TRUE)
                                               | wine$TotalSulfurDioxide_IMP > quantile(wine$TotalSulfurDioxide_IMP,c(.99), na.rm=TRUE)
                                               | is.na(wine$TotalSulfurDioxide_IMP),1,0))
wine$TotalSulfurDioxide_IMP[is.na(wine$TotalSulfurDioxide_IMP)] = median(wine$TotalSulfurDioxide_IMP, na.rm=TRUE)
wine$TotalSulfurDioxide_IMP = as.numeric(ifelse(wine$TotalSulfurDioxide_IMP < quantile(wine$TotalSulfurDioxide_IMP, c(.01)),
                                               quantile(wine$TotalSulfurDioxide_IMP,c(.01)), wine$TotalSulfurDioxide_IMP))
wine$TotalSulfurDioxide_IMP = as.numeric(ifelse(wine$TotalSulfurDioxide_IMP > quantile(wine$TotalSulfurDioxide_IMP, c(.99)),
                                               quantile(wine$TotalSulfurDioxide_IMP,c(.99)), wine$TotalSulfurDioxide_IMP))

#pH
wine$pH_FLAG = as.factor(ifelse(is.na(wine$pH),1,0))
wine$pH_IMP[is.na(wine$pH_IMP)] = median(wine$pH, na.rm = TRUE)

#Sulfates
wine$Sulphates_FLAG = as.factor(ifelse(is.na(wine$Sulphates),1,0))
wine$Sulphates_IMP[is.na(wine$Sulphates_IMP)] = median(wine$Sulphates, na.rm = TRUE)

#Alcohol
wine$Alcohol_FLAG = as.factor(ifelse(is.na(wine$Alcohol),1,0))
wine$Alcohol_IMP[is.na(wine$Alcohol_IMP)] = median(wine$Alcohol, na.rm = TRUE)

#Stars
wine$STARS_FLAG = as.factor(ifelse(is.na(wine$STARS),1,0))
wine$STARS_IMP = na.aggregate(wine$STARS_IMP, wine$LabelAppeal)

# Create Red/Wine Flag Variable
plot(wine$VolatileAcidity_IMP)

# Visualize volatile acidity
ggplot(data=wine, aes(wine$VolatileAcidity_IMP)) + 
  geom_histogram(binwidth =1, 
                 col="BLUE", 
                 aes(fill=..count..))+
  scale_fill_gradient("Count", low = "blue", high = "red")

summary(wine$VolatileAcidity_IMP)

# Make new indicator that indicates red vs white based on volatile acidity
wine$VolatileAcidity_IMP_REDFLAG = ifelse(wine$VolatileAcidity_IMP > mean(wine$VolatileAcidity_IMP),1,0)
wine$ResidualSugar_IMP_REDFLAG = ifelse(wine$ResidualSugar_IMP < mean(wine$ResidualSugar_IMP),1,0)
wine$TotalSulfurDioxide_IMP_REDFLAG = ifelse(wine$TotalSulfurDioxide_IMP < mean(wine$TotalSulfurDioxide_IMP),1,0)
wine$Density_IMP_REDFLAG = ifelse(wine$Density_IMP > mean(wine$Density_IMP),1,0)
wine$TallyUp = wine$VolatileAcidity_IMP_REDFLAG + wine$ResidualSugar_IMP_REDFLAG + wine$TotalSulfurDioxide_IMP_REDFLAG + wine$Density_IMP_REDFLAG
wine$Final_REDFLAG = ifelse(wine$TallyUp > mean(wine$TallyUp),1,0)

pairs(wine[,c("Final_REDFLAG","VolatileAcidity_IMP")])

plot(wine$VolatileAcidity_IMP,wine$TARGET)

# Add Target Flag for 0 sale scenarios
wine$TARGET_Flag = ifelse(wine$TARGET > 0,1,0)
wine$TARGET_AMT = wine$TARGET - 1
wine$TARGET_AMT = ifelse(wine$TARGET_Flag == 0,NA,wine$TARGET-1)

# Create interaction terms and imputed variables from input variables
wine$STARSxLabelAppeal_IMP = wine$STARS_IMP * wine$LabelAppeal_IMP
wine$STARSxAlcohol_IMP = wine$STARS_IMP * wine$Alcohol_IMP

# Create bins for 3 levels of Residual Sugar
wine$ResidualSugar_bin[wine$ResidualSugar_IMP < quantile(wine$ResidualSugar_IMP, c(.25))] = "Low Sweetness"
wine$ResidualSugar_bin[wine$ResidualSugar_IMP >= quantile(wine$ResidualSugar_IMP, c(.25)) & 
                         wine$ResidualSugar_IMP < quantile(wine$ResidualSugar_IMP, c(.75))] = "Medium Sweetness"
wine$ResidualSugar_bin[wine$ResidualSugar_IMP >= quantile(wine$ResidualSugar_IMP, c(.75))] = "High Sweetness"
wine$ResidualSugar_bin = factor(wine$ResidualSugar_bin, levels = c("Low Sweetness", "Medium Sweetness", "High Sweetness"))

# Create bins for 3 levels of acidity
wine$pH_IMP_bin[wine$pH_IMP <= 3] = "Low Acidity"
wine$pH_IMP_bin[wine$pH_IMP > 3 & wine$pH_IMP < 3.5] = "Medium Acidity"
wine$pH_IMP_bin[wine$pH_IMP >= 3.5] = "High Acidity"
wine$pH_IMP_bin = factor(wine$pH_IMP_bin, levels = c("Low Acidity", "Medium Acidity", "High Acidity"))

# Confirm no NAs for imputed variables
missmap(wine)
summary(wine)

```
### The same data preparation steps are performed on the testing dataset below

``` {r}
### Testing Data Set
# Create imputed variables for each column
wine_test$FixedAcidity_IMP = wine_test$FixedAcidity
wine_test$VolatileAcidity_IMP = wine_test$VolatileAcidity
wine_test$CitricAcid_IMP = wine_test$CitricAcid
wine_test$ResidualSugar_IMP = wine_test$ResidualSugar
wine_test$Chlorides_IMP = wine_test$Chlorides
wine_test$FreeSulfurDioxide_IMP = wine_test$FreeSulfurDioxide
wine_test$TotalSulfurDioxide_IMP = wine_test$TotalSulfurDioxide
wine_test$Density_IMP = wine_test$Density
wine_test$pH_IMP = wine_test$pH
wine_test$Sulphates_IMP = wine_test$Sulphates
wine_test$Alcohol_IMP = wine_test$Alcohol
wine_test$LabelAppeal_IMP = wine_test$LabelAppeal
wine_test$AcidIndex_IMP = wine_test$AcidIndex
wine_test$STARS_IMP = wine_test$STARS

### Replace missing NA values with median and outliers with 1st or 99th percentile of training data
### Create FLAG variables where replaced.

#ResidualSugar
wine_test$ResidualSugar_FLAG = as.factor(ifelse(wine_test$ResidualSugar_IMP < quantile(wine$ResidualSugar_IMP, c(.01), na.rm=TRUE)
                                           | wine_test$ResidualSugar_IMP > quantile(wine_test$ResidualSugar_IMP,c(.99), na.rm=TRUE)
                                           | is.na(wine_test$ResidualSugar_IMP),1,0))
wine_test$ResidualSugar_IMP[is.na(wine_test$ResidualSugar_IMP)] = median(wine$ResidualSugar_IMP, na.rm=TRUE)
wine_test$ResidualSugar_IMP = as.numeric(ifelse(wine_test$ResidualSugar_IMP < quantile(wine_test$ResidualSugar_IMP, c(.01)),
                                           quantile(wine$ResidualSugar_IMP,c(.01)), wine_test$ResidualSugar_IMP))
wine_test$ResidualSugar_IMP = as.numeric(ifelse(wine_test$ResidualSugar_IMP > quantile(wine_test$ResidualSugar_IMP, c(.99)),
                                           quantile(wine$ResidualSugar_IMP,c(.99)), wine_test$ResidualSugar_IMP))

#Chlorides
wine_test$Chlorides_FLAG = as.factor(ifelse(is.na(wine_test$Chlorides),1,0))
wine_test$Chlorides_IMP[is.na(wine_test$Chlorides_IMP)] = median(wine$Chlorides, na.rm = TRUE)

#FreeSulfurDioxide
wine_test$FreeSulfurDioxide_FLAG = as.factor(ifelse(wine_test$FreeSulfurDioxide_IMP < quantile(wine$FreeSulfurDioxide_IMP, c(.01), na.rm=TRUE)
                                               | wine_test$FreeSulfurDioxide_IMP > quantile(wine_test$FreeSulfurDioxide_IMP,c(.99), na.rm=TRUE)
                                               | is.na(wine_test$FreeSulfurDioxide_IMP),1,0))
wine_test$FreeSulfurDioxide_IMP[is.na(wine_test$FreeSulfurDioxide_IMP)] = median(wine$FreeSulfurDioxide_IMP, na.rm=TRUE)
wine_test$FreeSulfurDioxide_IMP = as.numeric(ifelse(wine_test$FreeSulfurDioxide_IMP < quantile(wine_test$FreeSulfurDioxide_IMP, c(.01)),
                                               quantile(wine$FreeSulfurDioxide_IMP,c(.01)), wine_test$FreeSulfurDioxide_IMP))
wine_test$FreeSulfurDioxide_IMP = as.numeric(ifelse(wine_test$FreeSulfurDioxide_IMP > quantile(wine_test$FreeSulfurDioxide_IMP, c(.99)),
                                               quantile(wine$FreeSulfurDioxide_IMP,c(.99)), wine_test$FreeSulfurDioxide_IMP))

#TotalSulfurDioxide
wine_test$TotalSulfurDioxide_FLAG = as.factor(ifelse(wine_test$TotalSulfurDioxide_IMP < quantile(wine$TotalSulfurDioxide_IMP, c(.01), na.rm=TRUE)
                                                | wine_test$TotalSulfurDioxide_IMP > quantile(wine_test$TotalSulfurDioxide_IMP,c(.99), na.rm=TRUE)
                                                | is.na(wine_test$TotalSulfurDioxide_IMP),1,0))
wine_test$TotalSulfurDioxide_IMP[is.na(wine_test$TotalSulfurDioxide_IMP)] = median(wine$TotalSulfurDioxide_IMP, na.rm=TRUE)
wine_test$TotalSulfurDioxide_IMP = as.numeric(ifelse(wine_test$TotalSulfurDioxide_IMP < quantile(wine_test$TotalSulfurDioxide_IMP, c(.01)),
                                                quantile(wine$TotalSulfurDioxide_IMP,c(.01)), wine_test$TotalSulfurDioxide_IMP))
wine_test$TotalSulfurDioxide_IMP = as.numeric(ifelse(wine_test$TotalSulfurDioxide_IMP > quantile(wine_test$TotalSulfurDioxide_IMP, c(.99)),
                                                quantile(wine$TotalSulfurDioxide_IMP,c(.99)), wine_test$TotalSulfurDioxide_IMP))

#pH
wine_test$pH_FLAG = as.factor(ifelse(is.na(wine_test$pH),1,0))
wine_test$pH_IMP[is.na(wine_test$pH_IMP)] = median(wine$pH, na.rm = TRUE)

#Sulfates
wine_test$Sulphates_FLAG = as.factor(ifelse(is.na(wine_test$Sulphates),1,0))
wine_test$Sulphates_IMP[is.na(wine_test$Sulphates_IMP)] = median(wine$Sulphates, na.rm = TRUE)

#Alcohol
wine_test$Alcohol_FLAG = as.factor(ifelse(is.na(wine_test$Alcohol),1,0))
wine_test$Alcohol_IMP[is.na(wine_test$Alcohol_IMP)] = median(wine$Alcohol, na.rm = TRUE)

#Stars
wine_test$STARS_FLAG = as.factor(ifelse(is.na(wine_test$STARS),1,0))
wine_test$STARS_IMP = na.aggregate(wine_test$STARS_IMP, wine_test$LabelAppeal)

# Make new indicator that indicates red vs white based on volatile acidity
wine_test$VolatileAcidity_IMP_REDFLAG = ifelse(wine_test$VolatileAcidity_IMP > mean(wine_test$VolatileAcidity_IMP),1,0)
wine_test$ResidualSugar_IMP_REDFLAG = ifelse(wine_test$ResidualSugar_IMP < mean(wine_test$ResidualSugar_IMP),1,0)
wine_test$TotalSulfurDioxide_IMP_REDFLAG = ifelse(wine_test$TotalSulfurDioxide_IMP < mean(wine_test$TotalSulfurDioxide_IMP),1,0)
wine_test$Density_IMP_REDFLAG = ifelse(wine_test$Density_IMP > mean(wine_test$Density_IMP),1,0)
wine_test$TallyUp = wine_test$VolatileAcidity_IMP_REDFLAG + wine_test$ResidualSugar_IMP_REDFLAG + wine_test$TotalSulfurDioxide_IMP_REDFLAG + wine_test$Density_IMP_REDFLAG
wine_test$Final_REDFLAG = ifelse(wine_test$TallyUp > mean(wine_test$TallyUp),1,0)

# Add Target Flag for 0 sale scenarios
wine_test$TARGET_Flag = ifelse(wine_test$TARGET > 0,1,0)
wine_test$TARGET_AMT = wine_test$TARGET - 1
wine_test$TARGET_AMT = ifelse(wine_test$TARGET_Flag == 0,NA,wine_test$TARGET-1)

# Create interaction terms and imputed variables from input variables
wine_test$STARSxLabelAppeal_IMP = wine_test$STARS_IMP * wine_test$LabelAppeal_IMP
wine_test$STARSxAlcohol_IMP = wine_test$STARS_IMP * wine_test$Alcohol_IMP

# Create bins for 3 levels of sweetness from the ResidualSugar
wine_test$ResidualSugar_bin[wine_test$ResidualSugar_IMP < quantile(wine$ResidualSugar_IMP, c(.25))] = "Low Sweetness"
wine_test$ResidualSugar_bin[wine_test$ResidualSugar_IMP >= quantile(wine$ResidualSugar_IMP, c(.25)) & 
                         wine_test$ResidualSugar_IMP < quantile(wine$ResidualSugar_IMP, c(.75))] = "Medium Sweetness"
wine_test$ResidualSugar_bin[wine_test$ResidualSugar_IMP >= quantile(wine$ResidualSugar_IMP, c(.75))] = "High Sweetness"
wine_test$ResidualSugar_bin = factor(wine_test$ResidualSugar_bin, levels = c("Low Sweetness", "Medium Sweetness", "High Sweetness"))

# Create bins for acidity based on pH level
wine_test$pH_IMP_bin[wine_test$pH_IMP <= 3] = "Low Acidity"
wine_test$pH_IMP_bin[wine_test$pH_IMP > 3 & wine_test$pH_IMP < 3.5] = "Medium Acidity"
wine_test$pH_IMP_bin[wine_test$pH_IMP >= 3.5] = "High Acidity"
wine_test$pH_IMP_bin = factor(wine_test$pH_IMP_bin, levels = c("Low Acidity", "Medium Acidity", "High Acidity"))

# Confirm there are no missing or empty values remaining in the dataset.
#summary(wine_test)
#missmap(wine_test)

```
### Part 4: Model Development
This section of the model is used to define and create 7 machine learning models. The models developed in this section are included below. This section begins with 2 linear regression models that are simple in nature to establish a baseline understanding of the data and relationships. The next five models compare to common Counting regression models, the Poisson and Negative Binomial regressions. These two models are then developed using the zero-inflated approach to control for the high volume of zeros in the dataset. Finally, a hurdle model is used to understand the probabilities of zero and non-zero target values.
1. Linear Regression
2. Stepwise Linear Regression
3. Poisson Regression
4. Negative Binomial Regression
5. Zero-Inflated Poisson (ZIP) Regression
6. Zero-Inflated Negative-Binomial (ZINB) Regression
7. Hurdle Model Regression
```{r}
### Model 1: Linear Regression
lm_fit = lm(TARGET~ ResidualSugar_bin + VolatileAcidity_IMP + CitricAcid_IMP +
              Chlorides_IMP + FreeSulfurDioxide_IMP + TotalSulfurDioxide_IMP + STARSxAlcohol_IMP +
              pH_IMP_bin + Sulphates_IMP + Alcohol_IMP + LabelAppeal_IMP + AcidIndex_IMP + STARS_IMP +
              Final_REDFLAG + STARSxLabelAppeal_IMP + ResidualSugar_FLAG + Chlorides_FLAG + FreeSulfurDioxide_FLAG +
              TotalSulfurDioxide_FLAG + pH_FLAG + Sulphates_FLAG + Alcohol_FLAG + STARS_FLAG
             ,data = wine)

summary(lm_fit)
wine$lm_fit = fitted(lm_fit)

### Model 2: Stepwise Linear Regression
stepwise_lm = stepAIC(lm_fit, direction="both", trace = 0)
stepwise_lm$anova

lm_fit_stepwise = lm(TARGET~ ResidualSugar_bin + VolatileAcidity_IMP + 
                       Chlorides_IMP + FreeSulfurDioxide_IMP + TotalSulfurDioxide_IMP +
                       pH_IMP_bin + Sulphates_IMP + STARSxAlcohol_IMP + LabelAppeal_IMP + AcidIndex_IMP + STARS_IMP +
                       Final_REDFLAG + STARSxLabelAppeal_IMP + ResidualSugar_FLAG
                       + pH_FLAG + STARS_FLAG
                       , data=wine)

summary(lm_fit_stepwise)
wine$fittedLMStepwise = fitted(lm_fit_stepwise)

### Model 3: Poisson Regression

poisson_model = glm(TARGET ~ ResidualSugar_bin + VolatileAcidity_IMP + 
                      Chlorides_IMP + FreeSulfurDioxide_IMP + TotalSulfurDioxide_IMP +
                      pH_IMP_bin + Sulphates_IMP + Alcohol_IMP + LabelAppeal_IMP + AcidIndex_IMP + STARS_IMP +
                      Final_REDFLAG + STARSxLabelAppeal_IMP + STARS_FLAG, 
                      family="poisson"(link="log"), data=wine)

summary(poisson_model)
wine$poisson_fit = predict(poisson_model, newdata = wine, type = "response")

### Model 4: Negative Binomial Regression

NBR_Model = glm.nb(TARGET ~ ResidualSugar_bin + VolatileAcidity_IMP + 
                     Chlorides_IMP + FreeSulfurDioxide_IMP + TotalSulfurDioxide_IMP +
                     pH_IMP_bin + Sulphates_IMP + STARSxAlcohol_IMP + LabelAppeal_IMP + AcidIndex_IMP + STARS_IMP +
                     Final_REDFLAG + STARSxLabelAppeal_IMP + ResidualSugar_FLAG +
                     pH_FLAG + STARS_FLAG, data=wine)

summary(NBR_Model)
wine$NBR_fit = predict(NBR_Model, newdata = wine, type = "response")

### Model 5: Zero-Inflated Poisson (ZIP) Regression

ZIP_Model = zeroinfl(TARGET ~ VolatileAcidity_IMP +
                      FreeSulfurDioxide_IMP + TotalSulfurDioxide_IMP +
                      pH_IMP_bin + Sulphates_IMP + STARSxAlcohol_IMP + LabelAppeal_IMP + AcidIndex_IMP +
                      Final_REDFLAG + STARSxLabelAppeal_IMP + STARS_IMP +
                      pH_FLAG + STARS_FLAG, data=wine)

summary(ZIP_Model)
wine$ZIP_fit = predict(ZIP_Model, newdata = wine, type = "response")

### Model 6: Zero-Inflated Negative Binomial (ZINB) Regression

ZINB_Model = zeroinfl(TARGET ~ VolatileAcidity_IMP + Alcohol_IMP +
                        FreeSulfurDioxide_IMP + TotalSulfurDioxide_IMP +
                        pH_IMP_bin + Sulphates_IMP + STARSxAlcohol_IMP + LabelAppeal_IMP + AcidIndex_IMP +
                        Final_REDFLAG + STARSxLabelAppeal_IMP + STARS_IMP +
                        pH_FLAG + STARS_FLAG, data=wine, dist = "negbin")

summary(ZINB_Model)
wine$ZINB_fit = predict(ZINB_Model, newdata = wine, type = "response")

### Model 7: Hurdle Model Regression
hurdle_model = hurdle(TARGET ~ VolatileAcidity_IMP + 
                        FreeSulfurDioxide_IMP + TotalSulfurDioxide_IMP +
                        pH_IMP_bin + Sulphates_IMP + STARSxAlcohol_IMP + LabelAppeal_IMP + AcidIndex_IMP +
                        Final_REDFLAG + STARSxLabelAppeal_IMP + STARS_IMP +
                        pH_FLAG + STARS_FLAG, data = wine)
summary(hurdle_model)
wine$hurdle_fit = predict(hurdle_model, newdata = wine, type = "response")

```
### Part 5: Model Evaluation
This section of the model is used to evaluate and compare the results of the ML regression models above. The models are compared using the following metrics:

1. Root Mean Squared Error (RMSE)
2. Mean Squared Error (MSE)
3. Akaike's Information Criteria (AIC)
4. Bayesian Information Criteria (BIC)

Additionally, the Vuong test is applied to compare the Poisson and Negative Binomial against their zero-inflated counterparts in order to establish statistical significance.
```{r, results = 'hide'}
# Define function for MSE calculation
mse = function(sm) 
  mean(sm$residuals^2)

# Define function for RMSE calculation
rmse = function(pred, target)
  sqrt(mean((pred-target)**2))

# AIC
lm_fit_AIC = AIC(lm_fit)
lm_fit_stepwise_AIC = AIC(lm_fit_stepwise)
poisson_model_AIC = AIC(poisson_model)
NBR_Model_AIC = AIC(NBR_Model)
ZIP_Model_AIC = AIC(ZIP_Model)
ZINB_Model_AIC = AIC(ZINB_Model)
hurdle_model_AIC = AIC(hurdle_model)

# BIC
lm_fit_BIC = BIC(lm_fit)
lm_fit_stepwise_BIC = BIC(lm_fit_stepwise)
poisson_model_BIC = BIC(poisson_model)
NBR_Model_BIC = BIC(NBR_Model)
ZIP_Model_BIC = BIC(ZIP_Model)
ZINB_Model_BIC = BIC(ZINB_Model)
hurdle_model_BIC = BIC(hurdle_model)

# MSE
lm_fit_mse = mse(lm_fit)
lm_fit_stepwise_mse = mse(lm_fit_stepwise)
poisson_model_mse = mse(poisson_model)
NBR_Model_mse = mse(NBR_Model)
ZIP_Model_mse = mse(ZIP_Model)
ZINB_Model_mse = mse(ZINB_Model)
hurdle_model_mse = mse(hurdle_model)

# RMSE
lm_fit_rmse = rmse(wine$lm_fit, wine$TARGET)
lm_fit_stepwise_rmse = rmse(wine$fittedLMStepwise, wine$TARGET)
poisson_model_rmse = rmse(wine$poisson_fit, wine$TARGET)
NBR_Model_rmse = rmse(wine$NBR_fit, wine$TARGET)
ZIP_Model_rmse = rmse(wine$ZIP_fit, wine$TARGET)
ZINB_Model_rmse = rmse(wine$ZINB_fit, wine$TARGET)
hurdle_model_rmse = rmse(wine$hurdle_fit, wine$TARGET)
```
```{r}
# Create table to display model evaluation results
data = matrix(c(lm_fit_AIC, lm_fit_stepwise_AIC, poisson_model_AIC, NBR_Model_AIC, ZIP_Model_AIC, ZINB_Model_AIC, hurdle_model_AIC, lm_fit_BIC, lm_fit_stepwise_BIC, poisson_model_BIC, NBR_Model_BIC, ZIP_Model_BIC, ZINB_Model_BIC, hurdle_model_BIC, lm_fit_mse, lm_fit_stepwise_mse, poisson_model_mse, NBR_Model_mse, ZIP_Model_mse, ZINB_Model_mse, hurdle_model_mse, lm_fit_rmse, lm_fit_stepwise_rmse, poisson_model_rmse, NBR_Model_rmse, ZIP_Model_rmse, ZINB_Model_rmse, hurdle_model_rmse), ncol = 7, byrow = TRUE)

# Add row and column names for results output
colnames(data) = c('LRM', 'LRM Stepwise','Poisson','Neg Binom', 'ZIP', 'ZINB', 'Hurdle')
rownames(data) = c('AIC', 'BIC', 'MSE', 'RMSE')
output = as.table(data)

print('The results of all models are compared in the table below: ')
print(output)

# Perform Vuong test to compare 2 poisson models
vuong(poisson_model, ZIP_Model)

# Perform Vuong test to compare 2 negative binomial models
vuong(NBR_Model, ZINB_Model)

# A summary of the 3 best-performing models that capture overdispered and zero-adjusted results
summary(wine$ZIP_fit)
summary(wine$ZINB_fit)
summary(wine$hurdle_fit)

```
### Part 6: Model Selection and Testing Prediction
The final portion of the script selects the highest-performing model and runs the prediction on the testing data set.
```{r}
### Highest-Performing Model: Zero-Inflated Negative Binomial Regression Model
# The results of the Vuong test demonstrated that the zero-inflated models scored much better compared to the model counterparts. Additionally, the ZINB model had the lowest RMSE and MSE compared to other 2 zero-inflated models. This model is output below and applied to the testing dataset. 
summary(ZINB_Model)

# Run prediction on testing dataset
wine_test$TARGET = predict(ZIP_Model, newdata = wine_test, type = "response")

summary(wine_test)
summary(wine_test$TARGET)
length(wine_test$TARGET[wine_test$TARGET<1]) / length(wine_test$TARGET)
summary(wine$TARGET)

```